\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=2.5cm]{geometry}
\usepackage{parskip}
\usepackage{amsmath,bm}

\title{Efficient Weight Computation}
\date{September 2018}
\author{Yu Guo, Milo\v s Ha\v san and Shuang Zhao}

\input{macros}

\begin{document}

\maketitle

This is a supplementary document to the paper ``Position-free Monte Carlo simulation for arbitrary layered BSDFs'', appearing in SIGGRAPH Asia 2018.

\section{Weights of Light Transport Paths}

Given a light subpath $\bar{x}_i$ and a camera subpath $\bar{x}_o$ with $n_i$ and $n_o$ vertices respectively, our bidirectional estimator combines $2 n_i n_o$ estimators of the form $f(\bar{y}^{(u)}_{s,t})/p^{(u)}_{s,t}(\bar{y}^{(u)}_{s,t})$ with $s \in \{1, 2, \ldots, n_i\}$, $t \in \{1, 2, \ldots, n_o\}$, and $u \in \{0, 1\}$ via the multiple importance sampling (MIS) framework.
This yields a combined estimator:
%
\begin{equation}
\label{eqn:bdpt_estimator}
\sum_{s = 1}^{n_i} \sum_{t = 1}^{n_o} \sum_{u = 0}^1 w^{(u)}_{s, t}(\bar{y}^{(u)}_{s,t}) \frac{f(\bar{y}^{(u)}_{s,t})}{p^{(u)}_{s,t}(\bar{y}^{(u)}_{s,t})},
\end{equation}
%
where the weight $w^{(u)}_{s, t}$, when using the balanced heuristics~[Veach 1997], is given by
%
\begin{equation}
\label{eqn:bdpt_weight_0}
w^{(u)}_{s, t}(\bar{y}_{s,t})
= \left(\sum_{s' = 1}^{s + t - 1} \sum_{u'=0}^1 \frac{p^{(u')}_{s', s + t - s'}(\bar{y}_{s,t})}{p^{(u)}_{s,t}(\bar{y}_{s,t})}\right)^{-1}
\end{equation}
%
for any path $\bar{y}_{s,t}$ with $(s + t)$ vertices.

Notice that, compared to standard bidirectional path tracing that combines $n_i n_o$ estimators, our position-free formulation offers twice the number of estimators since the direction connecting two depths is not unique (see Eq.~(16) of the main paper).


\section{Efficient Weight Computation}

Computing Eqs.~\eqref{eqn:bdpt_estimator} and \eqref{eqn:bdpt_weight_0} for all $s$ and $t$ na\"ively has a time complexity of $O(n_i n_o (n_i + n_o))$ and is too slow to be practical.
We now present our method that runs in $O(n_i n_o)$ time.
Our approach is conceptually similar to Veach's method for standard BDPT but differs in the exact mathematical form due to our position-free path formulation (see \S4 for the paper).

% \begin{comment}
% We start with rewriting Eq.~\eqref{eqn:bdpt_weight_0} as
% %
% \begin{equation}
% \label{eqn:bdpt_weight_1}
% w^{(u)}_{s, t}(\bar{y}) =  \left(1 + \sum_{s' = 1}^{s - 1} v_{0,s'} + \sum_{t' = 1}^{t - 1} v_{1,t'}\right)^{-1},
% \end{equation}
% %
% where
% %
% \begin{equation}
% v_{0,s'} := \sum_{u'=0}^1 \frac{p^{(u')}_{s', s + t - s'}(\bar{y})}{p^{(u)}_{s,t}(\bar{y})},
% \qquad
% v_{1,t'} := \sum_{u'=0}^1 \frac{p^{(u')}_{s + t - t', t'}(\bar{y})}{p^{(u)}_{s,t}(\bar{y})}.
% \end{equation}
% \end{comment}
%
Let $\bar{y}_{s,t} = (\bd_0, z_1, \bd_1, \ldots, z_n, \bd_n)$ with $n = s + t$.
For all $s', t' \in \{ 1, 2, \ldots, n \}$, define
%
\begin{align}
p^{(0)}_{s'} &:= \prod_{i = 1}^{s' - 1} p( \bd_i \;|\; z_i, \bd_{i - 1} ) \, p( z_{i + 1} \;|\; z_i, \bd_i ),\\
p^{(1)}_{t'} &:= \prod_{i = n - t' + 1}^{n - 1} p( -\bd_i \;|\; z_{i + 1}, -\bd_{i + 1} ) \, p( z_i \;|\; z_{i + 1}, -\bd_i ),
\end{align}
%
which denote the probability for constructing two subpaths containing the first $s'$ and last $t'$ vertices of $\bar{y}$, respectively.
Then, for all $u'$, $s'$ and $t'$, it holds that
%
\begin{equation}
p^{(u')}_{s',t'}(\bar{y}_{s,t}) = p^{(0)}_{s'} \, p^{(1)}_{t'} \, q^{(u')}_{s'},
\end{equation}
%
where
%
\begin{equation}
q^{(u')}_{s'} := \begin{cases}
p( \bd_{s'} \;|\; z_{s'}, \bd_{s' - 1} ) & \text{if $u = 0$,}\\
p( -\bd_{s'} \;|\; z_{s' + 1}, -\bd_{s' + 1} ) & \text{if $u = 1$.}
\end{cases}
\end{equation}
%
It follows that
%
\begin{equation}
\label{eqn:bdpt_pdf_ratio_0}
\frac{p^{(u')}_{s', t'}(\bar{y}_{s,t})}{p^{(u)}_{s,t}(\bar{y}_{s,t})}
= \frac{p^{(0)}_{s'} \, p^{(1)}_{t'} \, q^{(u')}_{s'}}{p^{(0)}_s \, p^{(1)}_t \, q^{(u)}_s}.
\end{equation}
%
Note that, for any $s' < s$, we have
%
\begin{equation}
p^{(u')}_{s', t'}(\bar{y}_{s,t}) = p^{(0)}_{s'} \, \frac{p^{(1)}_{t'}}{p^{(1)}_{t + 1}} \, q^{(u')}_{s'} \,p^{(1)}_{t + 1}.
\end{equation}
%
% \begin{comment}
% It follows that
% %
% \begin{equation}
% \label{eqn:bdpt_pdf_ratio_1}
% \frac{p^{(u')}_{s', t'}(\bar{y})}{p^{(u)}_{s,t}(\bar{y})}
% = \frac{p^{(0)}_{s'}(\bar{y})}{p^{(0)}_s(\bar{y})}
% \frac{p^{(1)}_{t'}(\bar{y})}{p^{(1)}_t(\bar{y})}
% \frac{q^{(u')}_{s'}(\bar{y})}{q^{(u)}_s(\bar{y})}.
% \end{equation}
% \end{comment}
%
It follows that
%
\begin{equation}
\label{eqn:bdpt_pdf_ratio_1}
\sum_{s' = 1}^{s - 1} \sum_{u' = 0}^1 \frac{p^{(u')}_{s', t'}(\bar{y}_{s,t})}{p^{(u)}_{s,t}(\bar{y}_{s,t})}
= \frac{p^{(1)}_{t + 1}}{p^{(1)}_t \, q^{(u)}_s}
\underbrace{\sum_{s' = 1}^{s - 1} \sum_{u' = 0}^1 \frac{p^{(0)}_{s'} \frac{p^{(1)}_{t'}}{p^{(1)}_{t + 1}} q^{(u')}_{s'}}{p^{(0)}_s}}_{=:\ P^{(0)}_s}.
\end{equation}
%
Since
%
\begin{equation}
\frac{p^{(1)}_{t'}}{p^{(1)}_{t + 1}}
= \prod_{i = s' + 1}^{s - 1} p( -\bd_i \;|\; z_{i + 1}, -\bd_{i + 1} ) \, p( z_i \;|\; z_{i + 1}, -\bd_i ),
\end{equation}
%
it is easy to verify that $P^{(0)}_s$ depends only on depths $z_{s'}$ and directions $\bd_{s'}$ with $s' \leq s$, which are all from the subpath $\bar{x}_i$.
Further, $P^{(0)}_{s'}$ remains constant for all paths $\bar{y}_{s,t}$ with $s > s'$.
This allows us to precompute $P^{(0)}_s$ using $\bar{x}_i$ for $s = 1, 2, \ldots, n_i$.
To this end, $P^{(0)}_s(\bar{y})$ can be efficiently evaluated using the following relation:
%
\begin{equation}
\label{eqn:bdpt_radio0}
P^{(0)}_s = \begin{cases}
0 & (s = 0),\\
\frac{p^{(0)}_{s - 1}}{p^{(0)}_s} \left( P^{(0)}_{s - 1} \frac{p^{(1)}_{t + 2}}{p^{(1)}_{t + 1}} + \sum_{u'} q^{(u')}_{s - 1} \right) & (s > 1).
\end{cases}
\end{equation}
%
Using Eq.~\eqref{eqn:bdpt_radio0}, we can compute $P^{(0)}_s(\bar{x}_i)$ for $s = 1, 2, \ldots, n_i$ in $O(n_i)$ time.

Similarly, for all $t' < t$, we have
%
\begin{equation}
\label{eqn:bdpt_pdf_ratio_2}
\sum_{t' = 1}^{t - 1} \sum_{u' = 0}^1 \frac{p^{(u')}_{s', t'}(\bar{y}_{s,t})}{p^{(u)}_{s,t}(\bar{y}_{s,t})}
= \frac{p^{(0)}_{s + 1}}{p^{(0)}_s \, q^{(u)}_s}
\underbrace{\sum_{t' = 1}^{t - 1} \sum_{u' = 0}^1 \frac{\frac{p^{(0)}_{s'}}{p^{(0)}_{s + 1}} p^{(1)}_{t'} q^{(u')}_{n - t'}}{p^{(1)}_t}}_{=:\ P^{(1)}_t},
\end{equation}
%
where $P^{(1)}_t$ only depends on $\bar{x}_o$ can be computed in $O(n_o)$ time.

With both $P^{(0)}_s$ and $P^{(1)}_t$ precomputed, Eq.~\eqref{eqn:bdpt_weight_0} becomes
%
\begin{equation}
\small
\label{eqn:bdpt_weight_1}
w^{(u)}_{s, t}(\bar{y}_{s,t})
= \left(1 + P^{(0)}_s + P^{(1)}_t +
\sum_{u'=0}^1 \frac{p^{(u')}_{s - 1, t + 1}(\bar{y}_{s,t}) + p^{(u')}_{s + 1, t - 1}(\bar{y}_{s,t})}{p^{(u)}_{s,t}(\bar{y}_{s,t})}\right)^{-1},
\end{equation}
%
which can be computed in constant time.
This leads to a full bidirectional estimator with time complexity $O(n_i n_o)$.

\end{document}
